### 2017-03-30 thesis/results/results.tex
tmp commit

\section{Obesity associated genetic signature from \citet{Creighton2012} study}
\label{sec:creighton_obesity_metagene}

@@

One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
Another reason for this apparent lack of association could be that the obesity associated genetic signature was too specific to the CR data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.
% Another reason for this apparent lack of association could be that the genetic signature was too specific to the data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the samples that were profiled in the Creighton \textit{et al.} publication.

\section{Obesity associated genetic signature from \citet{Fuentes-Mattei2014} study}
\label{sec:fm_obesity_metagene}

The obesity associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
Since the FM data set did not have patient \gls{bmi} information, the FM obesity metagene was not able to be compared with the patient \gls{bmi} or \gls{bmi} status in the original FM data.
Nevertheless, the transformation matrix was still generated in the FM data and applied first to the microarray data (CR and \gls{nzbc} data sets) then to the \gls{icgc} data sets to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in these data sets.
% However, the transformation matrix was generated in the FM data and applied to other cancer data sets  to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in those data sets.

@@

Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting FM metagene was compared with the sample gene expression, patient \gls{bmi} and \gls{bmi} status in each of the \gls{icgc} cancer data.
As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}; \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}).
The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value \cref{fig:fmmetaicgc}.
On the contrary to the association of the metagene with the patient \gls{bmi} status, the FM obesity metagene was not associated with the patient \gls{bmi}.
These results suggested that the samples from the patients that were overweight in the \gls{blca} data set had similar biological properties as the samples taken from the patients that were obese in the FM data set.
However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.

@@

\section{Novel obesity associated genetic signatures from \citet{Creighton2012} data set}
\label{sec:creighton_obesity_metagene_new}

@@

\begin{figure}[htpb]
	\centering
	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
	High and low correlation were represented as red and blue, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
	}
	\label{fig:cr_meta_cor}
\end{figure}

@@

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes with the patient \gls{bmi} and \gls{bmi} status in the \gls{icgc} \gls{blca} cancer data}
	\label{tab:degmetablca}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			 Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr & {\bfseries 0.0134}\tnote{1} & 0.1365 & {\bfseries 0.0387} & 0.0171 & {\bfseries 0.0195} \\
			Res                  & {\bfseries 0.0055}          & 0.1974 & {\bfseries 0.0196} & 0.0117 & {\bfseries 0.0446} \\
			CrOl                 & {\bfseries 0.0383}          & 0.4583 & 0.1096             & 0.0047 & 0.1374             \\
			ResOl                & 0.0909                      & 0.7092 & 0.2125             & 0.0018 & 0.2254             \\
			Ca                   & {\bfseries 0.0077}          & 0.1973 & {\bfseries 0.0231} & 0.0116 & {\bfseries 0.0456} \\
			CaRes                & {\bfseries 0.0104}          & 0.2712 & {\bfseries 0.0322} & 0.0101 & 0.0572             \\
			CaOl                 & 0.0575                      & 0.6185 & 0.1487             & 0.0022 & 0.2126             \\
			CaResOl              & {\bfseries 0.0463}          & 0.5820 & 0.1263             & 0.0036 & 0.1660             \\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\item [1] All values in bold are statistically significant (p \textless{} 0.05).
		\end{tablenotes}
	\end{threeparttable}
\end{table}

>>>>

\section{Obesity associated genetic signature from the\\\citet{Creighton2012} study}
\label{sec:creighton_obesity_metagene}

@@

One possible reason why the obesity metagene from the \citet{Creighton2012} study did not show significant association with other data sets could be because the genetic signature was in fact not an obesity specific signature, but a signature that was detected due to another clinical variable (investigated in \cref{sec:creighton_obesity_metagene_new}).
Another reason for this apparent lack of association could be that the obesity associated genetic signature was too specific to the CR data and was not a broad obesity associated genetic signature, but an obesity associated signature that was specific to the patient cohort that was profiled in the \citet{Creighton2012} publication.

\section{Obesity associated genetic signature from the\\\citet{Fuentes-Mattei2014} study}
\label{sec:fm_obesity_metagene}

The obesity associated genetic signature from the \citet{Creighton2012} study was not associated with patient \gls{bmi} or \gls{bmi} status in majority of the cancer data sets, so the obesity associated genetic signature from the \citet{Fuentes-Mattei2014} study (FM) was examined to see whether this obesity metagene was able to significantly associate with the patient \gls{bmi} and \gls{bmi} status across different  data sets.
Since the FM data set did not have patient \gls{bmi} information, the FM obesity metagene was not able to be compared with the patient \gls{bmi} or \gls{bmi} status in the original FM data.
Nevertheless, the transformation matrix was still generated in the FM data and applied first to the microarray data (CR and \gls{nzbc} data sets) then to the \gls{icgc} data sets to see whether the FM obesity metagene associated with patient \gls{bmi} and \gls{bmi} status in these data sets.

@@

Next, the transformation matrix was applied to the \gls{icgc} cancer data and the resulting FM metagene was compared with the sample gene expression, patient \gls{bmi} and \gls{bmi} status in each of the \gls{icgc} cancer data.
As evident in \cref{fig:fmmetaicgc,sec:rest_of_the_fm_icgc_cancer_heatmap_results}, the FM obesity metagene scores appeared to reflect the overall gene expression of the FM obesity associated genetic signature.
As with all the results in this chapter so far, the FM obesity metagene did not significantly associate with any of the \gls{icgc} cancer data, except in the \gls{blca} data set (\cref{fig:fmmetaicgc}; \cref{sec:rest_of_the_fm_icgc_cancer_heatmap_results}).
The FM obesity metagene significantly associated with the overweight group (but not with the obese group), and also had a significant \gls{anova} p-value \cref{fig:fmmetaicgc}.
On the contrary to the association of the metagene with the patient \gls{bmi} status, the FM obesity metagene was not associated with the patient \gls{bmi}.
These results suggested that the samples from the patients that were overweight in the \gls{blca} data set had similar biological properties as the samples taken from the patients that were obese in the FM data set.
However, due to the fact that the FM obesity metagene lacked association with the patient \gls{bmi} in the \gls{blca} data set, and that the metagene did not show any significant association in any of the other \gls{icgc} cancer types, it was difficult to determine whether the observed association of the FM obesity metagene with the overweight group was truly reflective of the quality of the FM metagene or observed by chance.

@@

\section{Novel obesity associated genetic signatures from the \citet{Creighton2012} data set}
\label{sec:creighton_obesity_metagene_new}

@@

\begin{figure}[tpb]
	\centering
	\includegraphics[width=0.8\linewidth]{results1/cr_meta_cor}
	\caption[Pearson correlation of all eight obesity metagenes identified in the CR data]{Heatmap showing the Pearson correlation of all eight obesity associated metagenes from the CR data with one another.
	\gls{svd} was applied to the \gls{rma} normalised CR data to generate each of the eight metagenes.
	High and low correlation were represented as red and yellow, respectively, where the colours were matched with the values on the scale shown in the top right histogram.
	The lowest correlation was 0.8469, between the CrOl and the CaRes metagenes.
	}
	\label{fig:cr_meta_cor}
\end{figure}

@@

\begin{table}[htpb]
	\centering
	\caption{Statistics of all the obesity metagenes with the patient \gls{bmi} and \gls{bmi} status in the \gls{icgc} \gls{blca} cancer data}
	\label{tab:degmetablca}
	\begin{threeparttable}
		\begin{tabular}{lccccc}
			& \multicolumn{3}{c}{ P-values} & \multicolumn{2}{c}{ Regression line statistics}\\
			\cmidrule(r){2-4} \cmidrule(r){5-6}
			 Metagenes &  Overweight &  Obese &  \gls{anova} &  R$^2$ &  P \\
			\hline
			\hline
			\rule{0pt}{2.25ex}Cr & {\bfseries 0.0134}\tnote{1} & 0.1365 & {\bfseries 0.0387} & 0.0171 & {\bfseries 0.0195} \\
			Res                  & {\bfseries 0.0055}          & 0.1974 & {\bfseries 0.0196} & 0.0117 & {\bfseries 0.0446} \\
			CrOl                 & {\bfseries 0.0383}          & 0.4583 & 0.1096             & 0.0047 & 0.1374             \\
			ResOl                & 0.0909                      & 0.7092 & 0.2125             & 0.0018 & 0.2254             \\
			Ca                   & {\bfseries 0.0077}          & 0.1973 & {\bfseries 0.0231} & 0.0116 & {\bfseries 0.0456} \\
			CaRes                & {\bfseries 0.0104}          & 0.2712 & {\bfseries 0.0322} & 0.0101 & 0.0572             \\
			CaOl                 & 0.0575                      & 0.6185 & 0.1487             & 0.0022 & 0.2126             \\
			CaResOl              & {\bfseries 0.0463}          & 0.5820 & 0.1263             & 0.0036 & 0.1660             \\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\begin{footnotesize}
				\item [1] All values in bold are statistically significant (p \textless{} 0.05).
			\end{footnotesize}
		\end{tablenotes}
	\end{threeparttable}
\end{table}

### 2017-03-30 thesis/results/results2.tex
pretty much ready to submit - tmp commit
tmp commit

\section{Pathway associated genetic signatures from \citet{Gatza2010a} study}
\label{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}

@@

Another thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data}).
This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
Taken together, it was decided that the GT data set should be batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).

@@

\begin{table}[htpb]
	\centering
	\begin{threeparttable}
	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets]{Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets}
	\label{tab:svd_vs_tm_path}
		\begin{tabular}{lcccc}
			& GT & CR & \gls{nzbc} & FM\\
			\hline
			\hline
			\rule{0pt}{2.25ex}Akt & 1.000 & 0.5190 & 0.5900 & 0.5563 \\
			\gls{bcat}            & 1.000 & 0.9897 & 0.9977 & 0.9905 \\
			E2F1                  & 1.000 & 0.9646 & 0.8438 & 0.9193 \\
			\gls{egfr}            & 1.000 & 0.0430 & 0.3358 & 0.4040 \\
			\gls{er}              & 1.000 & 0.9978 & 0.9942 & 0.9966 \\
			\gls{her2}            & 1.000 & 0.9553 & 0.5817 & 0.9794 \\
			\gls{ifna}            & 1.000 & 0.9830 & 0.9991 & 0.9951 \\
			\gls{ifny}            & 1.000 & 0.9086 & 0.9950 & 0.9718 \\
			Myc                   & 1.000 & 0.9878 & 0.9852 & 0.9689 \\
			p53                   & 1.000 & 0.3808 & 0.9981 & 0.7923 \\
			p63                   & 1.000 & 0.8319 & 0.2951 & 0.8368 \\
			\gls{pi3k}            & 1.000 & 0.9543 & 0.5989 & 0.9365 \\
			\gls{pr}              & 1.000 & 0.9511 & 0.9887 & 0.9845 \\
			Ras                   & 1.000 & 0.9078 & 0.9125 & 0.7229 \\
			Src                   & 1.000 & 0.9575 & 0.7173 & 0.9548 \\
			\gls{stat3}           & 1.000 & 0.1902 & 0.9159 & 0.6167 \\
			\gls{tgfb}            & 1.000 & 0.9918 & 0.2543 & 0.9890 \\
			\gls{tnfa}            & 1.000 & 0.6046 & 0.9365 & 0.4315 \\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\begin{footnotesize}
				\item [1] Transformation matrices were derived from the \gls{rma} normalised GT data set.
			\end{footnotesize}
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

\begin{table}[htpb]
	\centering
	\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the patient \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the patient \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
	\label{tab:lm_pr_only}
	\begin{threeparttable}
			\begin{tabular}{llrS}
				Linear Model & Variables & Estimate & {P-value}\\
				\hline
				\hline
				\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.477  & {\bfseries \num{6.06d-7}}\tnote{1}\\
				\hline
				\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002 & 0.586\\
                                                                           & \gls{pr}   & 0.478  & {\bfseries \num{6.35d-7}}\\
				\hline
				\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.085 & 0.176\\
                                                                           & Overweight & -0.011 & 0.875\\
                                                                           & \gls{pr}   & 0.471  & {\bfseries \num{8.25d-7}}\\
				\hline
				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007  & 0.237\\
                                                                           & Obese      & -0.188 & 0.081\\
                                                                           & Overweight & -0.049 & 0.516\\
                                                                           & \gls{pr}   & 0.459  & {\bfseries \num{1.54d-6}}\\
				\hline
				\hline
			\end{tabular}
			\begin{tablenotes}
				\begin{footnotesize}
					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
				\end{footnotesize}
			\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

Taken together, these results suggest that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
However, there were no common pathways that were present in all of the models; although Akt, \gls{egfr} and Src pathway metagenes were frequently included in many of the stepwise models.
The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.

>>>>

\section{Pathway associated genetic signatures from the\\\citet{Gatza2010a} study}
\label{sec:pathway_associated_genetic_signatures_from_gatza2010a_study}

@@

Another thing noted from these scatter plots was that there were some pathway signatures that were more variable than the others (\cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
As an example, the \gls{tgfb} metagenes were significantly more dispersed compared to the \gls{pr} metagenes, even though both of these metagenes were generated in the same data set through similar processes (\cref{fig:gt_rma_vs_mas}).
These differences were seen in other data sets as well (\cref{fig:appendix/gt_meta_rma_mas_other_data} in \cref{sec:normalisation_method_for_pathway_associated_genetic_signature_transformation_matrices}).
This result provided evidence that some of the pathway genetic signatures from the \citet{Gatza2010a} study were less variable across different data sets than the others.
Taken together, it was decided that the GT data set should be batch corrected first (\cref{sub:batch_correction}), then normalised with the \gls{rma} method, and the metagene scores were ranked with fractional ranking (\cref{sub:ranking_of_the_metagene_scores}).

@@

\begin{table}[tpb]
	\centering
	\begin{threeparttable}
	\caption[Summary of the Spearman correlations of the \gls{svd}- and TM-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets]{Summary of the Spearman correlations of the \gls{svd}- and transformation matrix\tnote{1}-derived pathway metagenes in the GT, CR, \gls{nzbc} and FM data sets}
	\label{tab:svd_vs_tm_path}
		\begin{tabular}{lcccc}
			& GT & CR & \gls{nzbc} & FM\\
			\hline
			\hline
			\rule{0pt}{2.25ex}Akt & 1.000 & 0.5190 & 0.5900 & 0.5563 \\
			\gls{bcat}            & 1.000 & 0.9897 & 0.9977 & 0.9905 \\
			E2F1                  & 1.000 & 0.9646 & 0.8438 & 0.9193 \\
			\gls{egfr}            & 1.000 & 0.0430 & 0.3358 & 0.4040 \\
			\gls{er}              & 1.000 & 0.9978 & 0.9942 & 0.9966 \\
			\gls{her2}            & 1.000 & 0.9553 & 0.5817 & 0.9794 \\
			\gls{ifna}            & 1.000 & 0.9830 & 0.9991 & 0.9951 \\
			\gls{ifny}            & 1.000 & 0.9086 & 0.9950 & 0.9718 \\
			Myc                   & 1.000 & 0.9878 & 0.9852 & 0.9689 \\
			p53                   & 1.000 & 0.3808 & 0.9981 & 0.7923 \\
			p63                   & 1.000 & 0.8319 & 0.2951 & 0.8368 \\
			\gls{pi3k}            & 1.000 & 0.9543 & 0.5989 & 0.9365 \\
			\gls{pr}              & 1.000 & 0.9511 & 0.9887 & 0.9845 \\
			Ras                   & 1.000 & 0.9078 & 0.9125 & 0.7229 \\
			Src                   & 1.000 & 0.9575 & 0.7173 & 0.9548 \\
			\gls{stat3}           & 1.000 & 0.1902 & 0.9159 & 0.6167 \\
			\gls{tgfb}            & 1.000 & 0.9918 & 0.2543 & 0.9890 \\
			\gls{tnfa}            & 1.000 & 0.6046 & 0.9365 & 0.4315 \\
			\hline
			\hline
		\end{tabular}
		\begin{tablenotes}
			\begin{footnotesize}
				\item [1] Transformation matrices were derived from the \gls{rma} normalised GT data set.
			\end{footnotesize}
		\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

\begin{table}[tpb]
	\centering
	\caption[Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the patient \gls{bmi}, \gls{bmi} status and the \acrshort{pr} pathway metagene score]{Description of the linear models constructed from the \gls{nzbc} data to predict the Cr obesity, using only the patient \gls{bmi}, \gls{bmi} status and the \gls{pr} pathway metagene score}
	\label{tab:lm_pr_only}
	\begin{threeparttable}
			\begin{tabular}{llrS}
				Linear Model & Variables & Estimate & {P-value}\\
				\hline
				\hline
				\rule{0pt}{2.25ex}\gls{pr} only                            & \gls{pr}   & 0.477  & {\bfseries \num{6.06d-7}}\tnote{1}\\
				\hline
				\rule{0pt}{2.25ex}\gls{bmi} and \gls{pr}                   & \gls{bmi}  & -0.002 & 0.586\\
                                                                           & \gls{pr}   & 0.478  & {\bfseries \num{6.35d-7}}\\
				\hline
				\rule{0pt}{2.25ex}\gls{bmi} status and \gls{pr}            & Obese      & -0.085 & 0.176\\
                                                                           & Overweight & -0.011 & 0.875\\
                                                                           & \gls{pr}   & 0.471  & {\bfseries \num{8.25d-7}}\\
				\hline
				\rule{0pt}{2.25ex}\gls{bmi}, \gls{bmi} status and \gls{pr} & \gls{bmi}  & 0.007  & 0.237\\
                                                                           & Obese      & -0.188 & 0.081\\
                                                                           & Overweight & -0.049 & 0.516\\
                                                                           & \gls{pr}   & 0.459  & {\bfseries \num{1.54d-6}}\\
				\hline
				\hline
			\end{tabular}
			\begin{tablenotes}
				\begin{footnotesize}
					\item [1] All values in bold are statistically significant (p \textless{} 0.05).
				\end{footnotesize}
			\end{tablenotes}
	\end{threeparttable}
\end{table}

@@

Taken together, these results suggest that the variables included in these stepwise models were relevant to the biology of the obesity metagenes identified in the CR and FM data sets.
However, there were no common pathways that were present in all of the models; although Akt, \gls{egfr}, \gls{tgfb} and Src pathway metagenes were frequently included in many of the stepwise models.
The role of the pathways included in the models and the possible biological links of these pathways to the obesity metagenes remain unclear, and should be the focus for future investigations.
